Extrinsic Calibration of 2D Laser Range Finders \\ using Planar Features

### 2019-03-26 paper.tex
changes after the review

% \author{\IEEEauthorblockN{Bernardo Lourenço}
% \IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
% \textit{Universidade de Aveiro}\\
% Aveiro, Portugal \\
% bernardo.lourenco@ua.pt}
% \and
% \IEEEauthorblockN{Paulo Dias}
% \IEEEauthorblockA{\textit{IEETA} \\
% \textit{Universidade de Aveiro}\\
% Aveiro, Portugal \\
% paulo.dias@ua.pt}
% \and
% \IEEEauthorblockN{Miguel Riem Oliveira}
% \IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
% \textit{Universidade de Aveiro}\\
% Aveiro, Portugal \\
% mriem@ua.pt}
% }

@@

Laser Range Finder, or LRFs, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{wang14,eyice18} and 3D reconstruction~\cite{saito10}. Their use is increasing due to several factors, such as decreasing costs, new research and development, and a fundamental role in emerging technologies such as autonomous driving.

@@

Therefore, reliable and automatic methods for the LRF extrinsic calibration are required. In this field, there are two main approaches, depending on the sensor which the LRF is calibrated relative to. The first method and the most studied is the calibration of the LRF relative to a camera~\cite{chen16,kassir10,zhang04}, or cameras~\cite{haeselich12}. In this method, the static transformation between these two sensors is determined and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF. This features might be markers placed on the field of view of both sensors, such as a chessboard, in the case of \cite{kassir10}. However, these methods have some disadvantages in the case of LRF mounted on robotic arms: first, they require the inclusion of a camera in the system; secondly, the calibration of the camera has to be performed; and lastly, the error of both calibrations is combined during the reconstruction, which increases the inaccuracies of the final reconstruction.

@@

The 2D-LRF is a sensor that outputs several laser scans. Each laser scan is a set of range measurements taken along several directions all coincident with a plane: the scanning plane. That is in fact why this equipment has a 2D nature. Thus, for each scan, a 2D slice of the scene is measured. Since the LRF is mounted on top of an actuated kinematic chain (the PTU), the movement of this chain positions the LRF in different poses in the 3D space. Through the accumulation of the several laser scans, a dense point cloud of the scene is produced.

@@

The segmentation was performed manually because it was not relevant for the scope of this work to perform it automatically. Also, it was a quick and one-time process, but it was necessary to ensure that the segmentation was well performed. In the future, automatic segmentation procedures could be studied to streamline the method.

@@

Further work can be done in order to improve or validate further this algorithm, such as: a study of a automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; and the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}.

>>>>

\author{\IEEEauthorblockN{Bernardo Lourenço}
\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
\textit{Universidade de Aveiro}\\
Aveiro, Portugal \\
bernardo.lourenco@ua.pt}
\and
\IEEEauthorblockN{Paulo Dias}
\IEEEauthorblockA{\textit{IEETA} \\
\textit{Universidade de Aveiro}\\
Aveiro, Portugal \\
paulo.dias@ua.pt}
\and
\IEEEauthorblockN{Miguel Riem Oliveira}
\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
\textit{Universidade de Aveiro}\\
Aveiro, Portugal \\
mriem@ua.pt}
}

@@

Laser Range Finder, or LRFs, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{cadena16} and 3D reconstruction~\cite{saito10}. Their use is increasing due to several factors, such as decreasing costs, new research and development, and a fundamental role in emerging technologies such as autonomous driving.

@@

Therefore, reliable and automatic methods for the LRF extrinsic calibration are required. In this field, there are two main approaches, depending on the sensor which the LRF is calibrated relative to. The first method and the most studied is the calibration of the LRF relative to a camera~\cite{chen16,vasconcelos12}, or cameras~\cite{haeselich12}. In this method, the static transformation between these two sensors is determined and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF. This features might be markers placed on the field of view of both sensors, such as a chessboard, in the case of \cite{kassir10}. However, these methods have some disadvantages in the case of LRF mounted on robotic arms: first, they require the inclusion of a camera in the system; secondly, the calibration of the camera has to be performed; and lastly, the error of both calibrations is combined during the reconstruction, which increases the inaccuracies of the final reconstruction.

@@

The 2D-LRF is a sensor that outputs several laser scans. Each laser scan is a set of range measurements taken sequentially along several directions all coincident with a plane: the scanning plane. That is in fact why this equipment has a 2D nature. Thus, for each scan, a 2D slice of the scene is measured. Since the LRF is mounted on top of an actuated kinematic chain (the PTU), the movement of this chain positions the LRF in different poses in the 3D space. Through the accumulation of the several laser scans, a dense point cloud of the scene is produced. It is important to consider that the scan is not acquired instantaneously. However, if the speed of the LRF is low, its effects are negligible.

@@

The segmentation was performed manually, using the software \textit{CloudCompare}\footnote{CloudCompare (\url{https://www.danielgm.net/cc/}) is a 3D point cloud and mesh processing software.}, because it was not relevant for the scope of this work to perform it automatically. Also, it was a quick and one-time process, but it was necessary to ensure that the segmentation was well performed. In the future, automatic segmentation procedures could be studied to streamline the method.

@@

Further work can be done in order to improve or validate further this algorithm, such as: a study of a automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}; and study the impact of the calibration environment (for example, the number and pose of the planes and size of the room) in the calibration accuracy.
